Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-23 ind_files/06_nontechnical_discussion.tex
Minor rephrasing

As explained in chapter \ref{chap_unification}, my general birth-death formalism reveals that the results of all these studies can be explained as specific instances of a more general phenomenon we call `noise-induced selection' (following previous studies such as~\cite{constable_demographic_2016} and~\cite{week_white_2021}). In particular, the equations of section \ref{sec_fun_theorems} generically predict a directional evolutionary force --- noise-induced selection --- acting on variation in per-capita turnover rate $\tau$ and favoring types whose per-capita turnover rate is lower than that of the population average. Noise-induced selection is only seen in finite populations, is seen whenever there is differential turnover rate $\tau$ in the system, and arises due to different types of individuals in the population experiencing a different number of stochastic events (birth and death) in a given time interval. This thesis is, to the best of my knowledge, the first time that noise-induced selection has been explicitly formulated and derived in full generality using a general `model-independent' evolutionary framework such as the Price equation (or my generalization thereof). My results thus unify several previous studies focused on specific systems under a single conceptual banner and show that their predictions are a generic outcome in finite, fluctuating populations of any nature as long as there is some heritable variation in per-capita turnover rates.

>>>>

\ref{chap_unification} formulates a set of stochastic differential equations derived from our general birth-death formalism which reveal that the results of all these studies can be explained as specific instances of a more general phenomenon we call `noise-induced selection' (following previous studies such as~\cite{constable_demographic_2016} and~\cite{week_white_2021}). In particular, the equations of section \ref{sec_fun_theorems} generically predict a directional evolutionary force --- noise-induced selection --- acting on variation in per-capita turnover rate $\tau$ and favoring types whose per-capita turnover rate is lower than that of the population average. Noise-induced selection is only seen in finite populations, is seen whenever there is differential turnover rate $\tau$ in the system, and arises due to different types of individuals in the population experiencing a different number of stochastic events (birth and death) in a given time interval. This thesis is, to the best of my knowledge, the first time that noise-induced selection has been explicitly formulated and derived in full generality using a general `model-independent' evolutionary framework such as the Price equation (or my generalization thereof). My results thus unify several previous studies focused on specific systems under a single conceptual banner and show that their predictions are a generic outcome in finite, fluctuating populations of any nature as long as there is some heritable variation in per-capita turnover rates.
